Wireless smart monitoring energy system: a proof of concept

### 2021-03-26 paper/main.tex
fdgbfdg

	\section{Introduction}
	Nowadays, keeping track of domestic energy usage is a common scenario in domotic applications that aim to monitor the real-time watt consumption of home appliances. In most cases, the solution is a sensor connected to the main safety switch yielding the total energy consumption. Our approach, to get more detailed information and to have control over the use of power by the appliances, is to have an energy sensor for each plug connected to the 220V line. In the following paragraphs, we present our \textit{proof of concept} (POC) which, functionally, consists of three main components: the smart plugs, the server, and the web interface. In our POC, we have modified the architecture described in \cite{8110428} obtaining a specific low-energy consumption-driven infrastructure. This consists of the smart plugs, a computer or a Raspberry Pi running the server, and the control devices which interact with the system through the web interface. In a household scenario, these devices are connected to the home wireless router. The communication between smart plugs and server occurs over UDP with a simple form of reliability implemented at the application level and tested using Wireshark \cite{wireshark}. The architecture implemented differs from the most common ones in two major aspects: first, most houses are not equipped with this kind of system, instead, a circuit breaker is triggered when too much power is being consumed. Second, when a similar system is used, it usually relies on cloud computing or external services. In contrast, our system offers the functionalities already described and both computing and data storage are performed locally. This can be an advantage in terms of both privacy and security. Furthermore, this is an advantage in terms of reliability: our system does not need an Internet connection to work properly, instead, it just needs a working WLAN. However, while this aspect is significant, the development of the prototype has not focused on the security aspects and some issues will be discussed in the appropriate section. The architecture of the system is such that if need be it can be easily modified to use third-party services. For example, consider the following use case: the user sets the maximum power consumption possible in the home and then turns on two appliances. When this happens for the first time the system decides, based on the type of appliance connected, a default value for the maximum power usage of the appliance. This value is then updated to reflect the real maximum power usage of the appliance. Now that the system is running and some appliances are turned on, the user will be able to switch on only the appliances that consume less than the available power. Our POC covers this use case and shows that such a system can be practical and useful and that the traffic generated is not enough to sensibly worsen the performances of the typical home WLAN. To summarize, the system keeps the power usage from reaching the maximum limit set by the user by forbidding the power-on of appliances that consume more than the available power avoiding blackouts. It also gives granular information about the energy consumption of the device and allows the user to turn on or off appliances remotely (via web interface). Aside from the main use case, in the age of climate awareness, such a system could be interesting to an environmentally conscious user that can monitor the power usage of appliances, but also could set the maximum power consumption to be lower than the real one to use less power and save money.

	\section{General architecture}
	The general architecture has three main components: the smart plugs, the server, and the web interface. The smart plugs code have been written using Arduino IDE in C++. Instead, the server has been developed in Java and the web interface is written in React. The rundown of what each part does is made up of a WeMos D1 R2, an Arduino-like board, monitors a sensor that gathers data on the power consumption, and controls a relay that turns on or off the appliance. It communicates over wireless with the UDP server that keeps track of the plugs and their energy consumption. The web interface uses an API exposed by the server to let the user see and control the status of each plug.
	\begin{figure}[htbp]
		\centering
		\includegraphics[width=\linewidth]{assets/architecture_schema}
		\caption{General architecture of the system}
		\label{fig:architecture_schema}
	\end{figure}

>>>>

	\section{Introduction}
	Nowadays, keeping track of domestic energy usage is a common scenario in domotic applications that aim to monitor the real-time watt consumption of home appliances. In most cases, the solution is a sensor connected to the main safety switch yielding the total energy consumption. Our approach, to get more detailed information and to have control over the use of power by the appliances, is to have an energy sensor for each plug connected to the 220V line. In the following paragraphs, we present our \textit{proof of concept} (POC) which, functionally, consists of three main components: the smart plugs, the server, and the web interface. In our POC, we have modified the architecture described in \cite{8110428} obtaining a specific low-energy consumption-driven infrastructure. This consists of the smart plugs, a computer or a Raspberry Pi running the server, and the control devices which interact with the system through the web interface. In a household scenario, these devices are connected to the home wireless router. The communication between smart plugs and server occurs over UDP with a simple form of reliability implemented at the application level and tested using Wireshark \cite{wireshark}. The architecture implemented differs from the most common ones in two major aspects: first, most houses are not equipped with this kind of system, instead, a circuit breaker is triggered when too much power is being consumed. Second, when a similar system is used, it usually relies on cloud computing or external services. In contrast, our system offers the functionalities already described and both computing and data storage are performed locally. This can be an advantage in terms of both privacy and security. Furthermore, this is an advantage in terms of reliability: our system does not need an Internet connection to work properly, instead, it just needs a working WLAN. However, while this aspect is significant, the development of the prototype has not focused on the security aspects and some issues will be discussed in the appropriate section. The architecture of the system is such that if need be it can be easily modified to use third-party services. For example, consider the following use case: the user sets the maximum power consumption possible in the home and then turns on two appliances. When this happens for the first time the system decides, based on the type of appliance connected, a default value for the maximum power usage of the appliance. This value is then updated to reflect the real maximum power usage of the appliance. Now that the system is running and some appliances are turned on, the user will be able to switch on only the appliances that consume less than the available power. Our POC covers this use case and shows that such a system can be practical and useful and that the traffic generated is not enough to sensibly worsen the performances of the typical home WLAN. To summarize, the system keeps the power usage from reaching the maximum limit set by the user by forbidding the power-on of appliances that consume more than the available power avoiding blackouts. It also gives granular information about the energy consumption of the device and allows the user to turn on or off appliances remotely (via web interface). Aside from the main use case, in the age of climate awareness, such a system could be interesting to an environmentally conscious user that can monitor the power usage of appliances, but also could set the maximum power consumption to be lower than the real one to use less power and save money. 
		\section{State of the Art}
	Smart plugs with power consumption detection are commonly used in the same household scenarios, but they implement different workflows and protocols. In fact, usually to realize these IoT devices an external platform is integrated with the main board (Arduino, Raspberry Pi, generic boards with EPS 8266) . In most cases, this platform is Blynk, which is a hardware-agnostic IoT platform and allows to configure our board in many different ways.
	The user could use the cloud infrastructure provided by the Blynk company, or, since the source code is open-source, download and install their own server. 
	This latter scenario is the preferred one since is free, and consist of the Blink server, the blink app and the smart devices, which uses Blynk API to perform the connections. A Blynk installation is generic-purpose, so you could install and control different types of devices and show the measurements using only the TCP protocols. Anyway, app user interface needs to be customized and the server needs a database, so the configuration requires much time. Instead, our system - as described in the following paragraphs - has a different goal and provides essential features to manage the smart plugs. The aim of our project is to create a functional IoT system which avoids blackouts and in the meantime has a low impact on the wireless network using a modified version of the UDP protocol.
	Although, Blynk implements some security feature which makes it more secure respect to our system.  
	\section{General architecture}
	The general architecture has three main components: the smart plugs, the server, and the web interface. The smart plugs code have been written using Arduino IDE in C++. Instead, the server has been developed in Java and the web interface is written in React. The rundown of what each part does is made up of a WeMos D1 R2, an Arduino-like board, monitors a sensor that gathers data on the power consumption, and controls a relay that turns on or off the appliance. It communicates over wireless with the UDP server that keeps track of the plugs and their energy consumption. The web interface uses an API exposed by the server to let the user see and control the status of each plug.
		\includegraphics[width=\linewidth]{assets/architecture_schema}
		\caption{General architecture of the system}
		\label{fig:architecture_schema}
